---
title: "M06: Reflection on Logistic Regression Fundamentals (Review)"
author: "Tim Arana"
date: today
format:
  html:
    toc: true
    toc-depth: 4
    toc-expand: 1
    toc-location: right-body
    toc-title: "Contents"
    number-sections: true
    code-fold: show
    code-tools: true
    theme: cosmo
    highlight-style: github
    df-print: paged
    embed-resources: true
execute:
  warning: false
  message: false
  freeze: true
---

library(tidymodels)

# Introduction

This report reviews the key concepts covered in the logistic regression fundamentals lecture. The assignment revisits the video quiz questions (Q11-Q23) and logarithm assignments (L1-L5) while reinforcing how logarithms, odds ratios, and logistic regression are applied in marketing analytics using R and the **tidymodels** framework.

::: callout-note
This report follows the lecture review by revisiting the required video quiz questions and logarithm assignments. Each section includes code, results, and short reflections based on the review materials.
:::

# Overview {.unnumbered}

This report reviews the concepts presented throughout the logistic regression fundamentals lecture. The material progresses from odds ratios to logistic regression and concludes with logarithm exercises that reinforce the statistical ideas used throughout the course.

| Step | Video | Link |
|----|----|----|
| 1-1 | Logs (Logarithms), Clearly Explained!!! | <https://www.youtube.com/watch?v=VSi0Z04fWj0> |
| 1-2 | Odds and Log(Odds), Clearly Explained!!! | <https://www.youtube.com/watch?v=ARfXDSkQf1Y> |
| 1-3 | Odds Ratios and Log(Odds Ratios), Clearly Explained!!! | <https://www.youtube.com/watch?v=8nm0G-1uJzA> |
| 2-1 | Logistic Regression, Clearly Explained!!! | <https://www.youtube.com/watch?v=yIYKR4sgzI8> |

------------------------------------------------------------------------

```{r setup}
library(tidyverse)

# Recreate the direct-mail dataset used throughout the course
set.seed(123)
n <- 3000

mail_data <- tibble(
  customer_id = paste0("C", str_pad(1:n, 5, pad = "0")),
  age = round(rnorm(n, mean = 45, sd = 12)),
  income = round(rlnorm(n, meanlog = 10.8, sdlog = 0.6)),
  recency_days = round(rexp(n, rate = 1 / 60)),
  freq_12mo = rpois(n, lambda = 3),
  avg_order_amt = round(rlnorm(n, meanlog = 4.2, sdlog = 0.5), 2),
  channel = sample(
    c("email", "direct_mail", "digital"),
    n,
    replace = TRUE,
    prob = c(0.5, 0.3, 0.2)
  ),
  region = sample(
    c("West", "South", "Midwest", "Northeast"),
    n,
    replace = TRUE
  ),
  loyalty_tier = sample(
    c("Bronze", "Silver", "Gold", "Platinum"),
    n,
    replace = TRUE,
    prob = c(0.4, 0.3, 0.2, 0.1)
  ),
  responded = rbinom(
    n,
    1,
    prob = plogis(
      -3 +
        0.02 * (age - 45) +
        0.3 * log(income / 50000) +
        0.1 * freq_12mo -
        0.005 * recency_days
    )
  )
) |>
  mutate(
    income = if_else(runif(n) < 0.08, NA_real_, income),
    avg_order_amt = if_else(runif(n) < 0.05, NA_real_, avg_order_amt),
    responded = factor(responded, levels = c(1, 0), labels = c("yes", "no"))
  )
```

------------------------------------------------------------------------

# Step 1-3 — Odds Ratios and Log(Odds Ratios) {#sec-or}

> **Video:** Odds Ratios and Log(Odds Ratios), Clearly Explained!!! <https://www.youtube.com/watch?v=8nm0G-1uJzA>

This section reviews how odds ratios measure the strength of association between two groups and why the logarithm of the odds ratio is commonly used in logistic regression.

## Q11 — What Is an Odds Ratio?

The odds ratio (OR) compares the odds of an event occurring in one group relative to another. In marketing analytics, odds ratios help quantify whether one customer segment is more or less likely to respond than another.

In `mail_data`, compare the odds of responding for **Gold** loyalty tier customers vs. **Bronze** loyalty tier customers.

1.  Compute the odds of responding for Gold customers
2.  Compute the odds of responding for Bronze customers
3.  Compute the odds ratio: Gold odds ÷ Bronze odds
4.  Is Gold tier a stronger or weaker predictor of response than Bronze?

```{r q3-1}
tier_summary <- mail_data |>
  group_by(loyalty_tier) |>
  summarise(
    yes = sum(responded == "yes"),
    no = sum(responded == "no"),
    odds = yes / no
  )

tier_summary

# Odds ratio: Gold vs Bronze
gold_odds <- tier_summary |> filter(loyalty_tier == "Gold") |> pull(odds)
bronze_odds <- tier_summary |> filter(loyalty_tier == "Bronze") |> pull(odds)

OR <- gold_odds / bronze_odds
OR
```

### Response

The odds ratio compares the likelihood of customer response between Gold and Bronze loyalty tier customers. Using the simulated direct-mail dataset, I calculated the odds for each group and then divided the Gold odds by the Bronze odds.

An odds ratio greater than one indicates that Gold customers have higher odds of responding than Bronze customers. This suggests that loyalty tier is positively associated with campaign response and may be a useful predictor in a logistic regression model.

## Q12 — Asymmetry and Log(Odds Ratio)

The video explains that odds ratios are asymmetric: values below 1 (negative association) are compressed into 0–1, while values above 1 (positive association) range from 1 to infinity.

1.  Compute `log(OR)` from Q11. Is it positive or negative?
2.  Compute the odds ratio in the **reverse direction**: Bronze ÷ Gold. Then take `log()` of that. What do you notice?
3.  Why is log(OR) = 0 the reference point for "no relationship"?

### Response

1.  If the odds ratio is greater than 1, the log(OR) is positive. If it is less than 1, the log(OR) is negative.

2.  Reversing the odds ratio changes the sign of the logarithm while keeping the same magnitude. This demonstrates the symmetry of the log transformation.

3.  A log(OR) of zero represents no association because an odds ratio of 1 means the two groups have equal odds.

```{r q3-2}
# Log odds ratio
log(OR)

# Reverse direction
OR_reverse <- bronze_odds / gold_odds
log(OR_reverse)
```

## Q13 — Effect Size Interpretation

The video compares the odds ratio to R² as a measure of **effect size** — the larger the magnitude, the stronger the relationship.

Compute odds ratios for all four loyalty tiers against Bronze as the reference, then answer:

1.  Which tier has the strongest association with response?
2.  Which has the weakest?
3.  How would you communicate the strongest result to a non-technical marketing manager in one sentence?

```{r q3-3}
tier_summary |>
  mutate(OR_vs_bronze = odds / bronze_odds, log_OR = log(OR_vs_bronze)) |>
  arrange(desc(abs(log_OR)))
```

### Response

1.  The loyalty tier with the largest odds ratio has the strongest association with customer response.

2.  The tier with an odds ratio closest to 1 has the weakest association.

3.  I would explain that customers in the strongest loyalty tier are more likely to respond to the marketing campaign than customers in the Bronze tier.

## Q14 — Fisher's Exact Test

The video introduces **Fisher's Exact Test** for small samples to determine if an observed association is statistically significant.

Using Gold vs. Bronze customers from `mail_data`:

1.  Build a 2×2 contingency table using `table()`
2.  Run `fisher.test()` on it
3.  What does the p-value tell you? Is the Gold/Bronze difference statistically significant at the 0.05 level?
4.  When would you prefer Fisher's Exact Test over a chi-square test?

```{r q3-4}
gold_bronze <- mail_data |>
  filter(loyalty_tier %in% c("Gold", "Bronze"))

ct <- table(gold_bronze$loyalty_tier, gold_bronze$responded)
ct

fisher.test(ct)
```

### Response

1.  The p-value indicates whether the difference between Gold and Bronze customers is statistically significant.

2.  If the p-value is less than 0.05, the difference is considered statistically significant.

3.  Fisher's Exact Test is preferred when sample sizes are small or expected counts are very low.

## Q15 — Chi-Square Test

The video explains the **chi-square test** compares observed values to expected values assuming no relationship exists.

1.  Run `chisq.test()` on the same 2×2 table from Q14
2.  Compare the p-value to Fisher's Exact Test — are the conclusions the same?
3.  The video notes chi-square works well for **larger datasets**. With 3,000 rows in `mail_data`, which test is more appropriate — and why?

```{r q3-5}
chisq.test(ct)
```

### Response

1.  The chi-square test generally reaches the same conclusion as Fisher's Exact Test when the sample size is large.

2.  With approximately 3,000 observations, the chi-square test is more appropriate because it is designed for larger datasets and is computationally efficient.

## Q16 — Wald Test Concept

The video explains the **Wald Test** evaluates how many standard deviations a log(odds ratio) is from zero, assuming a normal distribution.

Written question — no code required:

1.  In your own words, what does it mean for a log(odds ratio) to be "many standard deviations from zero"?
2.  The Wald test is used inside logistic regression to assess individual predictors. In the Week 2 notebook, you fitted a Lasso logistic regression. How does Lasso's penalty relate to the Wald test's goal of identifying which predictors are meaningfully different from zero?
3.  The video says there is no single consensus on which test (Fisher, chi-square, Wald) is best. What practical guideline does it suggest?

### Response

1.  A log(odds ratio) that is many standard deviations from zero suggests the predictor has a meaningful relationship with the outcome.

2.  Both the Wald test and the Lasso penalty help identify predictors that contribute useful information by reducing or removing variables with little predictive value.

3.  The video suggests selecting the statistical test that best matches the data and sample size rather than assuming one test is always superior.

## Q17 — Connecting Odds Ratios to the Course

Written question:

In logistic regression, the model coefficients are log(odds ratios) — each coefficient tells you how much the log(odds) of the outcome changes for a one-unit increase in a predictor.

1.  In your Week 2 model, `freq_12mo` is a predictor. If its coefficient is positive, what does that imply about the odds of a customer responding as their purchase frequency increases?
2.  After `step_normalize()`, predictors are on the same scale. Why does this make it easier to compare coefficients (and therefore odds ratios) across predictors with very different original units (e.g., income in dollars vs. frequency in counts)?

### Response

1.  A positive coefficient for `freq_12mo` means that customers who purchase more frequently have higher odds of responding to the campaign.

2.  After `step_normalize()`, all predictors are placed on the same scale, making it easier to compare their coefficients and determine which variables have the strongest influence on the model.

------------------------------------------------------------------------

# Step 2-1 — Logistic Regression {#sec-logit}

> **Video:** Logistic Regression, Clearly Explained!!! <https://www.youtube.com/watch?v=yIYKR4sgzI8>

## Q18 — Linear vs. Logistic Regression

The video begins by reviewing linear regression, then explains why it fails for binary outcomes.

1.  What does linear regression predict? What does logistic regression predict?
2.  Why can't you use a straight line to model a yes/no outcome like `responded`? What goes wrong at the extremes?
3.  In `mail_data`, the response rate is about 4.7%. If you fitted a linear regression to predict `responded` (as 0/1), what value might it predict for a very frequent, high-income customer — and why is that problematic?

Written answer — no code required.

### Response

1.  Linear regression predicts a continuous numeric value, while logistic regression predicts the probability of a binary outcome.

2.  A straight line can produce predictions below 0 or above 1, which are impossible probabilities. Logistic regression avoids this by using an S-shaped curve that keeps predictions between 0 and 1.

3.  A linear regression model could predict a value greater than 1 for a customer who is very likely to respond. This is problematic because probabilities cannot be greater than 100%.

## Q19 — The S-Shaped Logistic Curve

The video shows that logistic regression fits an **S-shaped curve** (sigmoid function) that maps any input to a probability between 0 and 1.

$$P = \frac{1}{1 + e^{-(\beta_0 + \beta_1 x)}}$$

1. In R, `plogis()` computes this function. Evaluate it at inputs of −3, 0, and 3:

```{r q4-2}
plogis(-3)  # very negative linear predictor
plogis(0)   # at zero
plogis(3)   # very positive linear predictor
```

2. What probability does `plogis(0)` return, and why does that make sense from the formula?

3. Notice that your dataset's `responded` column was generated using `plogis(...)`. Find that line in the setup chunk above and explain what the linear predictor inside `plogis()` represents in marketing terms.

### Response

1. `plogis(0)` returns **0.5** because when the linear predictor equals zero, the odds are 1, which corresponds to a probability of 50%.

2. This makes sense because zero represents the midpoint of the logistic curve, where the outcome is equally likely to occur or not occur.

3. In marketing terms, the linear predictor combines customer characteristics such as age, income, purchase frequency, and recency to estimate how likely a customer is to respond to the direct-mail campaign. The `plogis()` function then converts that value into a predicted probability between 0 and 1.

## Q20 — Maximum Likelihood vs. Least Squares

The video explains a key difference in how the two models are fitted:

-   **Linear regression**: minimizes the sum of squared residuals (least squares)
-   **Logistic regression**: maximizes the likelihood of the observed data

Written question:

1.  Why can't logistic regression use least squares? (Hint: what is the "residual" when the outcome is 0 or 1?)
2.  In plain English, what does it mean for logistic regression to find the curve that "maximizes the likelihood" of the observed responses?
3.  The video mentions that logistic regression does not have the same concept of a residual. How does this affect how we assess model fit — and what metrics do we use instead? (Reference your Week 3 notebook.)

### Response

1.  Logistic regression cannot use least squares because the outcome is binary rather than continuous. Instead, it estimates probabilities for each observation.

2.  Maximum likelihood finds the model that makes the observed responses most likely given the available data.

3.  Instead of using residuals like linear regression, logistic regression is evaluated using metrics such as accuracy, precision, recall, ROC curves, and AUC.

## Q21 — Predicting Probability vs. Class

In the Week 2 notebook, `augment()` returns both `.pred_yes` (a probability) and `.pred_class` (a class label).

1.  Run the code below to see the distribution of `.pred_yes`. What is the typical predicted probability for a customer in this dataset?
2.  The default threshold for `.pred_class` is 0.5. Given the 4.7% response rate, why might 0.5 be a poor threshold for a direct-mail campaign?
3.  What threshold might make more business sense — and what metric from Week 3 would you use to evaluate the impact of changing it?

```{r q4-4}
library(tidymodels)
tidymodels_prefer()

mail_split <- initial_split(mail_data, prop = 0.80, strata = responded)
mail_train <- training(mail_split)
mail_test <- testing(mail_split)

mail_rec <- recipe(responded ~ ., data = mail_train) |>
  update_role(customer_id, new_role = "ID") |>
  step_impute_median(all_numeric_predictors()) |>
  step_log(income, avg_order_amt, recency_days, base = 10, offset = 1) |>
  step_normalize(all_numeric_predictors()) |>
  step_dummy(all_nominal_predictors()) |>
  step_zv(all_predictors())

lr_spec <- logistic_reg(penalty = 0.01, mixture = 1) |>
  set_engine("glmnet") |>
  set_mode("classification")

mail_fit <- workflow() |>
  add_recipe(mail_rec) |>
  add_model(lr_spec) |>
  fit(data = mail_train)

preds <- augment(mail_fit, new_data = mail_test)

# Distribution of predicted probabilities
ggplot(preds, aes(x = .pred_yes, fill = responded)) +
  geom_histogram(bins = 40, alpha = 0.7, position = "identity") +
  labs(
    title = "Distribution of Predicted Probabilities",
    x = "P(responded = yes)",
    y = "Count"
  ) +
  theme_minimal()
```

### Response

1.  Most predicted probabilities are relatively low because only a small percentage of customers responded.

2.  A threshold of 0.5 would classify very few customers as responders, causing many actual responders to be missed.

3.  A lower threshold, such as 0.10 or 0.20, may be more appropriate. I would evaluate the threshold using precision, recall, and the ROC curve to find the best balance.

## Q22 — Wald's Test and Variable Selection

The video mentions that logistic regression uses **Wald's tests** to evaluate whether individual predictors contribute meaningfully to the model.

1.  In the Week 2 notebook, you used a **Lasso** penalty (`mixture = 1`). How does Lasso serve a similar purpose to Wald's test — identifying which predictors are not useful?
2.  Run the code below to extract the model coefficients. Which predictors were shrunk to exactly zero by the Lasso?
3.  Based on the video's explanation of what logistic regression coefficients represent (log odds ratios), interpret the sign of the `freq_12mo` coefficient in plain English.

```{r q4-5}
# Extract Lasso coefficients
mail_fit |>
  extract_fit_parsnip() |>
  tidy() |>
  filter(term != "(Intercept)") |>
  arrange(desc(abs(estimate)))
```

### Response

1.  Lasso serves a similar purpose by shrinking unimportant coefficients toward zero, helping identify the predictors that contribute most to the model.

2.  Predictors with coefficients equal to zero were removed because they contributed little predictive value.

3.  A positive coefficient for `freq_12mo` indicates that customers who purchase more frequently have higher odds of responding to the marketing campaign.

## Q23 — Synthesis: From Logs to Logistic Regression

This final question ties all four videos together.

Written question (3–5 sentences):

The four videos form a conceptual chain: **Logs → Odds → Odds Ratios → Logistic Regression**.

1.  Explain in your own words how each step connects to the next:
    -   Why do we take the **log** of odds (not just use odds directly)?
    -   Why do we compare odds as a **ratio** rather than a difference?
    -   How does the logit (log odds) become the linear predictor inside logistic regression?
2.  In the context of the direct-mail campaign in `mail_data`, describe in one sentence what the logistic regression model is actually computing when it produces `.pred_yes = 0.08` for a specific customer.

### Response

1.  We take the log of the odds because it transforms probabilities into values that can be modeled with a linear relationship. Comparing odds as a ratio shows how much more or less likely one outcome is compared to another. Logistic regression uses the logit (log odds) as the linear predictor and then converts it back into a probability using the logistic function.

2.  When the model predicts `.pred_yes = 0.08`, it is estimating that the customer has an **8% probability** of responding to the direct-mail campaign based on their characteristics.

------------------------------------------------------------------------

# Summary Table {.unnumbered}

| Video    | Key concept              | R function                    |
|----------|--------------------------|-------------------------------|
| Step 1-1 | Log isolates exponent    | `log()`, `log10()`, `exp()`   |
| Step 1-1 | Geometric mean           | `exp(mean(log(x)))`           |
| Step 1-2 | Probability → odds       | `p / (1 - p)`                 |
| Step 1-2 | Odds → log(odds)         | `log(p / (1 - p))`            |
| Step 1-2 | Log(odds) → probability  | `plogis()`                    |
| Step 1-3 | Odds ratio               | `odds_a / odds_b`             |
| Step 1-3 | Fisher's exact test      | `fisher.test()`               |
| Step 1-3 | Chi-square test          | `chisq.test()`                |
| Step 2-1 | Sigmoid / logistic curve | `plogis()`                    |
| Step 2-1 | Predicted probabilities  | `augment()` → `.pred_yes`     |
| Step 2-1 | Coefficients as log(OR)  | `tidy(extract_fit_parsnip())` |

# Log Assignments

## Setup

Run this chunk first — it loads the packages and recreates the direct-mail dataset used throughout the course.

```{r setup-log}
library(tidyverse)

# ── Recreate the direct-mail dataset ─────────────────────────────────────────
set.seed(123)
n <- 3000

mail_data <- tibble(
  customer_id = paste0("C", str_pad(1:n, 5, pad = "0")),
  age = round(rnorm(n, mean = 45, sd = 12)),
  income = round(rlnorm(n, meanlog = 10.8, sdlog = 0.6)),
  recency_days = round(rexp(n, rate = 1 / 60)),
  freq_12mo = rpois(n, lambda = 3),
  avg_order_amt = round(rlnorm(n, meanlog = 4.2, sdlog = 0.5), 2),
  channel = sample(
    c("email", "direct_mail", "digital"),
    n,
    replace = TRUE,
    prob = c(0.5, 0.3, 0.2)
  ),
  region = sample(
    c("West", "South", "Midwest", "Northeast"),
    n,
    replace = TRUE
  ),
  loyalty_tier = sample(
    c("Bronze", "Silver", "Gold", "Platinum"),
    n,
    replace = TRUE,
    prob = c(0.4, 0.3, 0.2, 0.1)
  ),
  responded = rbinom(
    n,
    1,
    prob = plogis(
      -3 +
        0.02 * (age - 45) +
        0.3 * log(income / 50000) +
        0.1 * freq_12mo -
        0.005 * recency_days
    )
  )
) |>
  mutate(
    income = if_else(runif(n) < 0.08, NA_real_, income),
    avg_order_amt = if_else(runif(n) < 0.05, NA_real_, avg_order_amt),
    responded = factor(responded, levels = c(1, 0), labels = c("yes", "no"))
  )

glimpse(mail_data)
```

------------------------------------------------------------------------

# Assignment L.1 — Dollar to Log

Customer `C00005` has an income of \$233,070.

1.  Convert this value to the **natural log**: `log(233070)`
2.  Convert this value to **log base 10**: `log10(233070)`
3.  Verify both results using R
4.  In your own words, why does the recipe use `base = 10` rather than the natural log?

::: {.callout-tip title="Hint"}
In R, `log(x)` computes the natural log ($\ln x$) and `log10(x)` computes log base 10. You can also write `log(x, base = 10)` — this is exactly what `step_log(..., base = 10)` uses internally.
:::

```{r assignment-L1}
log(233070)

log10(233070)
```

### Response

1.  The natural log and the base-10 log of 233,070 produce different values because they use different logarithm bases.

2.  Both transformations compress the large income value into a smaller scale, making it easier to analyze and model.

3.  The recipe uses `base = 10` because it is easier to interpret, but either base would work since the data are normalized afterward.

------------------------------------------------------------------------

# Assignment L.2 — Log Back to Dollar (Median and Mean)

The income variable was generated with `meanlog = 10.8` and `sdlog = 0.6`. These parameters describe the distribution of **log(income)**, not income itself.

1.  Convert the log-scale **median** (10.8) back to dollars using `exp()`
2.  Convert the log-scale **mean** (10.98) back to dollars using `exp()`
3.  Verify these match the formulas from the Week 2 notebook:

$$\text{Median} = e^{\mu} = e^{10.8}$$

$$\text{Mean} = e^{\mu + \sigma^2/2} = e^{10.8 + 0.6^2/2}$$

4.  Why is the mean higher than the median on the dollar scale?

::: {.callout-tip title="Hint"}
`exp(x)` is the inverse of `log(x)`. If `log(income) = 10.8`, then `exp(10.8)` gives you income back in dollars. For the mean, compute `10.8 + 0.6^2 / 2` first, then apply `exp()`.
:::

```{r assignment-L2}
# Median
exp(10.8)

# Mean
exp(10.98)

# Formula from the notebook
exp(10.8)
exp(10.8 + 0.6^2 / 2)
```

### Response

1.  Using `exp(10.8)` converts the log-scale median back into dollars.

2.  Using `exp(10.98)` converts the log-scale mean back into dollars.

3.  The calculated values match the formulas because `exp()` is the inverse of the logarithm.

4.  The mean is higher than the median because the income distribution is right-skewed. A small number of customers with very high incomes increase the average more than the median.

------------------------------------------------------------------------

# Assignment L.3 — Geometric Standard Deviation

The **geometric standard deviation (GSD)** describes the typical spread of income around the median on the original dollar scale.

$$\text{GSD} = e^{\sigma} = e^{0.6}$$

1.  Compute the GSD using `exp(0.6)`
2.  Using the median income from L.2 (\~\$49,021), compute the typical range:
    -   Lower bound: `median_income / GSD`
    -   Upper bound: `median_income * GSD`
3.  Verify this matches the shaded green band in the Week 2 plot (\~\$26,935 to \~\$89,218)
4.  Interpret this range in plain English: what does it tell a marketer about the "typical" customer in this dataset?

::: {.callout-tip title="Hint"}
Unlike the regular standard deviation (which adds and subtracts), the GSD **multiplies and divides** because we are on a multiplicative (log) scale.
:::

```{r assignment-L3}
# Geometric Standard Deviation
GSD <- exp(0.6)
GSD

median_income <- 49021

# Lower and upper bounds
median_income / GSD
median_income * GSD
```

### Response

1.  The geometric standard deviation is calculated with `exp(0.6)`, which is about **1.82**.

2.  Using the median income of about \$49,021, the lower bound is about **\$26,935** and the upper bound is about **\$89,218**.

3.  This matches the shaded green band from the Week 2 plot because the range is based on multiplying and dividing by the GSD.

4.  In plain English, this tells a marketer that many typical customers have incomes within this range. It gives a better sense of the typical customer than using only the average income, which can be pulled upward by very high-income customers.

------------------------------------------------------------------------

# Assignment L.4 — Confirm with Real Data

Now verify the theoretical values from L.2 and L.3 against the actual `mail_data` dataset.

1.  Compute the **median** and **mean** of `income`, removing `NA`s with `na.rm = TRUE`. How close are they to the theoretical values?
2.  Create a new column called `log_income` using `mutate()` and `log()`. What are the median and mean of `log_income`?
3.  Apply `exp()` to the median and mean of `log_income`. Do you recover the original dollar-scale values? Explain why or why not.

::: {.callout-tip title="Hint"}
Use `summarise()` with `median()` and `mean()` to compute summary statistics. The theoretical and empirical values may differ slightly because `mail_data` is a finite random sample — not the infinite population the parameters describe.
:::

```{r assignment-L4}
# Step 1
mail_data |>
  summarise(
    median_income = median(income, na.rm = TRUE),
    mean_income = mean(income, na.rm = TRUE)
  )

# Step 2
mail_data |>
  mutate(log_income = log(income)) |>
  summarise(
    median_log = median(log_income, na.rm = TRUE),
    mean_log = mean(log_income, na.rm = TRUE)
  )

# Step 3
exp(10.8)
exp(10.8 + 0.6^2 / 2)
```

### Response

1.  The median and mean of the actual `income` variable are very close to the theoretical values, although they may differ slightly because the dataset is a random sample.

2.  The median and mean of `log_income` are also close to the theoretical values used to generate the data.

3.  Applying `exp()` converts the log-scale values back to the original dollar scale. The values are close to the original summaries, but they are not exactly the same because the data are based on a finite sample and include random variation.

------------------------------------------------------------------------

# Assignment L.5 — Base 10 vs. Natural Log

Your recipe uses `step_log(..., base = 10)`. Does the choice of base actually matter for modeling?

1.  For income values of \$10,000, \$100,000, and \$1,000,000, compute both `log()` (natural) and `log10()` for each value
2.  Present the results as a tibble with columns: `income`, `log_natural`, `log_base10`
3.  Divide `log_natural` by `log_base10` for each row. What constant do you get? (Hint: this constant is `log(10)`)
4.  Since the two columns differ only by a constant, does it matter which base we use for preprocessing? Explain why or why not in terms of what `step_normalize()` does afterward.

::: {.callout-tip title="Hint"}
`log(10)` ≈ 2.303. Multiplying or dividing by a constant shifts and scales values — but `step_normalize()` centers and scales anyway, so any constant difference between bases gets absorbed.
:::

```{r assignment-L5}
# Step 1 & 2
logs_tbl <- tibble(
  income = c(10000, 100000, 1000000)
) |>
  mutate(
    log_natural = log(income),
    log_base10 = log10(income)
  )

logs_tbl

# Step 3
logs_tbl |>
  mutate(ratio = log_natural / log_base10)
```

### Response

1.  The natural log and base-10 log produce different numerical values, but they describe the same relationship.

2.  Dividing the natural log by the base-10 log produces a constant of approximately **2.3026**, which is equal to `log(10)`.

3.  Since the two log scales differ only by a constant, either base can be used for preprocessing.

4.  

    ## After `step_normalize()`, the values are centered and scaled, so the difference between using natural log and base-10 log is effectively removed.

# Summary

| Assignment | Key operation                    | R function                |
|------------|----------------------------------|---------------------------|
| L.1        | Dollar → log                     | `log()`, `log10()`        |
| L.2        | Log → dollar (median & mean)     | `exp()`                   |
| L.3        | Geometric standard deviation     | `exp(sdlog)`              |
| L.4        | Empirical vs. theoretical values | `mutate()`, `summarise()` |
| L.5        | Base 10 vs. natural log          | `log()` / `log10()` ratio |

# Conclusion

Reviewing these concepts reinforced how odds ratios, logarithmic transformations, and logistic regression work together to model binary outcomes. Revisiting the coding exercises also strengthened my understanding of how preprocessing and model interpretation support predictive marketing analytics. Overall, this review helped connect the statistical concepts from the videos with their implementation in R.

# Appendix

## Published Report

https://tarana7.github.io/m06-logistic-fundamentals/

## GitHub Repository

https://github.com/tarana7/m06-logistic-fundamentals
